---
title: "gt_to_xlsx walkthrough"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(gtopenxlsx)
```

`gt_to_openxlsx` takes the great table as input. The first thing it will do, it will create an object call `ordered_data` (equivalent I guess of body in `gt`). All the following functions will then use the gt table and the ordered_data doing different manipulation.

There are 3 types of functions:  

- create_ordered_data: it will create the data.frame with the data.
- apply_*: these functions will apply formatting, merge.
- write_*: these functions will write the different part of the table in the sheet.

```{r, message=FALSE}
library(gtopenxlsx)
library(gt)
library(dplyr)
order_countries <- c("Germany", "Italy", "United States", "Japan")

gt_table <-
  gtcars |>
  arrange(
    factor(ctry_origin, levels = order_countries),
    mfr, desc(msrp)
  ) |>
  mutate(car = paste(mfr, model)) |>
  select(-mfr, -model) |>
  group_by(ctry_origin) |>
  gt(rowname_col = "car") |>
  cols_hide(columns = c(drivetrain, bdy_style)) |>
  cols_move(
    columns = c(trsmn, mpg_c, mpg_h),
    after = trim
  ) |>
  tab_spanner(
    label = "Performance",
    columns = c(mpg_c, mpg_h, hp, hp_rpm, trq, trq_rpm)
  ) |>
  cols_merge(
    columns = c(mpg_c, mpg_h),
    pattern = "<<{1}c<br>{2}h>>"
  ) |>
  cols_merge(
    columns = c(hp, hp_rpm),
    pattern = "{1}<br>@{2}rpm"
  ) |>
  cols_merge(
    columns = c(trq, trq_rpm),
    pattern = "{1}<br>@{2}rpm"
  ) |>
  cols_label(
    mpg_c = "MPG",
    hp = "HP",
    trq = "Torque",
    year = "Year",
    trim = "Trim",
    trsmn = "Transmission",
    msrp = "MSRP"
  ) |>
  fmt_currency(columns = msrp, decimals = 0) |> 
  tab_header(
    title = md("The Cars of **gtcars**"),
    subtitle = "These are some fine automobiles"
  ) |>
  tab_source_note(
    source_note = md(
      "Source: Various pages within the Edmonds website."
    )
  ) 

```

```{r, eval=FALSE}
library(openxlsx)
wb <- createWorkbook()
addWorksheet(wb, "gt_cars_example", gridLines = F)
sheet_name <-"gt_cars_example"

```

```{r, eval=FALSE}
 ordered_data <- gt_table |>
    create_ordered_data()

  if("_formats" %in% names( gt_table)) {
    ordered_data <- apply_formats(gt_table,
                                  ordered_data)

  }

  if("_col_merge" %in% names(gt_table)) {
    ordered_data <- apply_col_merge(gt_table,ordered_data)
  }

  current_row <- 1

  write_table_header(gt_table = gt_table,
                     ordered_gt_data = ordered_data,
                     wb = wb,
                     sheet_name = sheet_name,
                     row_to_start = current_row)
  if(!is.null(gt_table$`_heading`$title)) {
    current_row <- current_row + 1
  }

  if(!is.null(gt_table$`_heading`$subtitle)) {
    current_row <- current_row + 1
  }

  if (nrow(gt_table$`_spanners`) > 0) {
    write_spanners(gt_table = gt_table,
                   ordered_gt_data = ordered_data,
                   wb = wb,
                   sheet_name = sheet_name,
                   row_to_start = current_row)

    number_of_level <- gt_table$`_spanners`$spanner_level|> max()
    current_row <- current_row + number_of_level

  }


  write_stub_head_and_column_labels(gt_table = gt_table,
                                    ordered_gt_data = ordered_data,
                                    wb = wb,
                                    sheet_name = sheet_name,
                                    row_to_start = current_row)

  current_row <- current_row + 1

  list_cell_indexes <- write_stub_and_table_body(gt_table = gt_table,
             ordered_gt_data = ordered_data,
             wb = wb,
             sheet_name = sheet_name,
             row_to_start = current_row)


  if (length(gt_table$`_source_notes`) > 0) {
    final_index_stub <- list_cell_indexes |>
      purrr::map(~.x[["end"]]) |>
      purrr::list_c() |>
      max()

    current_row <- final_index_stub + 1
    write_source_note(gt_table = gt_table,
                      ordered_gt_data = ordered_data,
                      wb = wb,
                      sheet_name = sheet_name,
                      row_to_start = current_row)
  }
```

```{r,eval=FALSE}
saveWorkbook(wb, "gt_cars_example.xlsx")

```

